Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2025-06-27 Latex_files/Thesis/chapters/chapter 1/main.tex
27/06/2025 chapters 1,3,4 and structural changes

\chapter{Introduction}
Crystallisation is one of the oldest forms of phase 
separation used by humanity \cite{Schoen1956}, put 
simply, it is the formation and growth of a new 
structured phase within a disordered bulk phase. 
This has applications in a number of industries such 
as pharmaceuticals \cite{Gao2017}, food production 
\cite{Hartel2002}, and electronics \cite{Myerson2002}. 
Where the extraction of dilute materials can help 
improve product quality while keeping production costs 
low. The industrialisation of crystallisation has 
allowed engineers to reliably and efficiently induce 
the crystal formation within a bulk phase. However, 
while on a large scale crystallisation is seemingly an 
understood physical process, just a small amount of 
investigation into the literature reveals that at a 
micro scale there is yet to be unifying theory that can 
accurately explain the process of crystallisation 
\cite{Fu2021}. Crystallisation can be subdivided into 
two steps: Nucleation and Crystal growth. The latter 
focusing on how an already stable crystal grows and how 
it takes on its final shape. Whereas the former is more 
concerned with the first initial moments of a crystals 
formation within a bulk phase.

In this chapter we will outline our current understanding 
of nucleation, why there is a gap in the literature, and why 
there is still a need for local control of nucleation events. 
Furthermore, we will also highlight recent developments 
involving the use of optical tweezers (and other laser based 
methods) in order to develop our understanding of nucleation. 
Lastly we highlight the connection between optical tweezer's 
ability to probe fluid viscosity via rotational motion, and 
the known link between nucleation and fluid flow, to propose 
a novel potential method of creating localised nucleation events.

\section{Nucleation}
Nucleation is an example of a binary phase separation, 
where a dilute phase is miscible in a bulk phase, more 
often called the solute and solvent respectively. When 
a small amount of solute is added to the solvent the 
solute will diffuse throughout the solvent, reaching
an equilibrium concentration throughout the solution. 
Because of molecular interactions, the two can only remain 
in equilibrium while below a specific concentration ($C_{eq}$) 
- below which the chemical potential $\mu$ for a single phase 
is greater than the potential required to separate the two. 
Once $C_{eq}$ is exceeded there is a chemical potential 
difference driving the solution to separate the two phases. 
Since different combinations of solute and solvent will 
have different equilibrium concentrations, researchers often 
instead measure the ratio between the solute and solvent by 
using 'supersaturation' \cite{Mullin2001}:
\begin{align}
	\label{eq:supersaturation}
	S(T) = \frac{C_{sol}}{C_{eq}(T)}
\end{align}

Where $C_{sol}$ is just the solute concentration, and 
$C_{eq}(T)$ is the equilibrium concentration of the solute  
at temperature $T$ based on the main solvent. While the 
solution remains supersaturated there is a chemical 
potential driving the solute to coalesce and separate 
from the solution as an ordered solid, the first 
formation of the crystal is referred to as the 
nucleus and understanding its formation has been the 
focus of researchers for decades now. Typically, for an 
industrial crystallisation process the working principle 
is based on controlling and manipulating the supersaturation 
of the system.

\subsection{Primary \& Secondary nucleation}
From an industrial perspective, the nucleation process can 
be broadly categorised into either primary of secondary
nucleation. The former describes the formation of an initial
nucleus within the bulk phase, absent of any external stimuli.
Primary nucleation is therefore considered stochastic as there 
is no means of predicting where a nucleus may form, or how long 
it will take. The only reliable indicator being that higher 
supersaturations will result in faster nucleation rates. At a 
small scale one can estimate the nucleation rate by making 
repeated measurements of sample solutions and seeing how many 
have nucleated after a given time, giving us a Poisson probability distribution.
\begin{align}
	P(t) = 1 - exp\left[-JV(t-t_g)\right] = \frac{M^*(t)}{M}
	\label{eq:poisson}
\end{align}

Where $J$ is the nucleation rate, $t_g$ is the 'growth time', 
$V$ is the volume of the individual samples, and $M^*(t)$ \& 
$M$ are the number of nucleated samples and the total number 
of samples used respectively. The nucleation rate is one of 
the most important factors used in evaluating a crystallisers 
performance, providing a rough estimation of the final product 
mass over a given time frame. While \eqref{eq:poisson} is useful 
for studying the effects of different parameters at a small scale, 
for industrial applications there are to many external factors 
for primary nucleation to be measured accurately.

Secondary nucleation is the result of a initial seed crystal 
inducing further nucleation within the bulk solution 
\cite{Botsaris1976}. The seed crystals is often specially 
prepared ahead of time and added to a supersaturate solution. 
Due to interactions with the surrounding fluid and the container 
walls it is possible for the seed crystals to act as a surface
for further nucleation \cite{Anwar2015}. Control of secondary 
nucleation events are crucial for ensuring reliable industrial 
crystalliser performance. Fig.~\ref{fig:secondary} depicts an 
attempt of classifying every possible mechanism that could lead 
to secondary nucleation.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.95\linewidth]{secondary_nucleation.jpg}
	\caption{Secondary Nucleation mechanisms, classified by
		Agrawal and Paterson \cite{Agrawal2015}}
	\label{fig:secondary}
\end{figure}

This is not a universal classification system, there are a 
variety of opinions on how best to characterise different 
phenomena. It is heavily dependent on the theory used to 
describe nucleation events, this is accurate for both 
secondary and primary nucleation.

\section{Nucleation Theories}
In order to predict the nucleation rate of a given solution
we need a solid understanding of how the individual molecules
interact such that they transition from an unordered phase to
a ordered phase. Over the years several theories have been 
proposed to explain how nucleation events occur, so far there
has yet to be a standout candidate that reliably predicts 
the systems behaviour. Either they fail to accurately predict
nucleation rates or more generally cannot describe new 
experimental results.

\subsection{Classical Nucleation Theory (CNT)}
Sometimes referred to as 'Gibbs Nucleation Theory' the 
original theory was first formed from the works of Volmer 
and Weber, and Frenkel \cite{Frenkel1939, Volmer1926}. 
While initially it was more focused on describing droplet
formation in condensing vapours it was extrapolated to 
describe crystallisation. The central premise of classical 
theory is that nucleation occurs stochastically due to 
collisions between individual solute molecules, ions, or 
atoms. At the same time the bulk phase is resistant to the
formation of a new phase. The competition between these 
random collisions and the bulk solution can be used to 
predict the probability of a newly formed nucleus.

Consider a supersaturated solution, at some time individual 
sub units collide rapidly, one after another, forming a 
nucleus of volume $4\pi r^3/3$. The newly formed phase has a 
lower chemical potential than the surrounding solution, 
reducing the free energy of the system. Simultaneously, 
the formation of a new interface is resited by the bulk 
phase due to surface tension. The net free energy of the 
system for a nucleus of radius $r$ is given as \cite{Karthika2016}:
\begin{align}
	\Delta G = \frac{-4\pi r^3}{3v}k_BT\ln(S) + 4\pi r^{2}\sigma_{inf}
	\label{eq:CNT} 
\end{align}

Where $S$ is the supersaturation from eq.~\eqref{eq:supersaturation} 
$v$ is the approximate volume of an individual molecule, 
$k_B$ is the Boltzmann constant, and $\sigma_{inf}$ is the 
interfacial tension of the bulk solution. This assumes that 
the nucleus will have a spherical morphology so that the 
surface tension $\sigma_{inf}$ is a scalar value. Looking at 
\eqref{eq:CNT} suggests that there must be some critical size 
$r$ where the free energy gain from the nucleus exceeds the 
surface tension of the surrounding fluid. This is reflected in 
fig.~\ref{fig:free_energy} where we plot the free energy of the 
system against nucleus size. This reveals a critical size above 
which the gain in free energy exceeds the interfacial tension. 
Furthermore fig.~\ref{fig:free_energy} shows how increasing the 
supersaturation of the system reduces said barrier. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{Free_Energy_Diagram.png}
	\caption{Free energy diagram of a newly formed nucleus according 
		     to the Classical Nucleation Theory. The total free energy (blue)
		     is due to the competition between the volume free energy gain
		     (green) and the interfacial free energy cost (orange). Dotted
		     lines are for a higher supersaturation than the solid lines,
		     the interfacial energy cost is independent of supersaturation.}
	\label{fig:free_energy}
\end{figure}

The maximum value of $\Delta G_{tot}$ is the free energy barrier 
that any newly formed nucleus needs to overcome in order to stabilise. 
The nucleation rate (the volume of new crystalline material formed per 
unit time), is therefore commonly defined as being dependent on the 
energy barrier $\Delta G^*$:
\begin{align}
	J = A \ exp \left[-\frac{\Delta G^*}{k_BT} \right]
\end{align}

Where $A$ is a pre-factor that can be fine tuned to the exact 
demands of the system, the free energy barrier can be found by 
finding the stationary point of $\Delta G_{tot}$.

CNT is often regarded as a good description of the macro system, 
its obvious that for all crystallization systems there is an 
inherent energy barrier that dictates the nucleation rate. 
Where it falters is in its predictive ability, both in estimating 
nucleation rates \cite{Gharibeh2005, Vekilov2010}, and in the 
structure of newly formed nuclei \cite{Lee1999, Yau2001}.
Recent studies suggest classical nucleation is merely one of 
many possible pathways that can be taken to produce an ordered 
crystalline phase. Prompting the development of alternative 
theories to better describe the nucleation process.

\subsection{Two Step Nucleation}
The two step nucleation theory is an extension to the CNT 
that suggests that prior to nucleation, the solute will 
form precursor structures. The CNT assumes that solute 
structures below the critical radius should be unstable 
within the solution due to interfacial forces trying to 
dissolve the solute. This assumption is based on the idea 
that any interactions between solute molecules are much
weaker than that between solute and solvent molecules.
If instead the solute molecule interactions are significant
then it stands to reason that subcritical liquid-like 
structures are stable.

The first observation of a liquid-like structure was in 
$CaCO_3$ solutions, initially it was assumed that polymers
were a necessary additive to create a precursor structure 
\cite{Driessche2017, Karthika2016}. This is not the case when 
the solution is above its critical temperature, allowing the 
solution to be in the binodal region of the phase diagram. 
This would result in the formation of two liquid phases, 
one being rich in solute, the other being solute poor 
\cite{Karthika2016, Fu2021}. Several papers later reported 
the presence of stable liquid-like clusters that formed 
prior to nucleation \cite{Savage2009, Wolde1997, Soga1999}.

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{oswalds_rule.jpg}
	\caption{Free energy diagram of Oswald's rule for a 
	crystal with three possible polymorphs. The diagram 
	shows that there exist local minima in the free energy 
	that represent the different polymorphic forms. A similar
	free energy diagram can be used to describe the two-step
	nucleation mechanism.}
\end{figure}

The formation of these clusters can be understood by Oswald's 
rule; which says that any crystallising system does not 
immediately take the path to the lowest possible energy state 
but instead first transitions to the state with the smallest 
free energy barrier \cite{Ostwald1897}. Oswald's rule was 
assumed to be applicable for crystal polymorphs but can be
generalised to describe the nucleation process also. Further phase 
transitions can still occur but the pathway taken should 
always minimises the overall free energy cost. It is assumed
that between the amorphous phase and the crystalline phase 
the solute undergoes dehydration while changing structurally 
into a more ordered phase, though there is some discussion 
whether the process merely undergoes dehydration, or if there 
is a more complex chemical reaction
\cite{Karthika2016}.

With regards to liquid-precursors, there is still a lot of
discussion about the relationship between the solution 
conditions and the the precursors. The size and rate at 
which these precursors agglomerate based on the degree of 
saturation would provide invaluable information on the 
kinetics of two-step nucleation \cite{Fu2021}. In some 
cases it has even been shown that 'ageing' solutions is 
a necessary step in order to detect the presence of 
stable amorphous clusters \cite{Liao2022}. This was in 
solutions that were supersaturated and yet after the 
formation of precursors spontaneous nucleation was not
seen even after several days of observation \cite{Liao2022}.
This suggests that in some cases the precursors are
metastable and require an external trigger to induce
crystallisation. The presence of metastable structures
and other phenomena has lead to development of a more 
universal theory, namely non-classical nucleation.

\subsection{Non-classical Nucleation} 
Two-step nucleation is based on the idea that between
the bulk solution and the final crystalline phase there
exists and intermediate stage. While this is broadly an
accepted concept, research has shown that in some cases
the intermediate stage can be subdivided into multiple 
states. This has been shown to be the case with metals
in both solid and liquid conditions \cite{Cao2020, Ye2023}
and colloidal simulations have indicated that the 
precursors undergo structural change over time \cite{Tan2013}. 
While in some cases it has been possible to directly
monitor the structural change in the amorphous clusters, 
many organic compounds have not received such treatment
due to the difficulty of characterising and monitoring 
the pre nucleation stage \cite{Fu2021}. This is often 
why two-step nucleation is simply referred to as 
multi-step nucleation.

There are also a myriad of other unique nucleation events
that deviate significantly from both the classical and 
two-step nucleation theories. A review paper discussing
recent advances in non-classical crystallisation highlighted
that depending on the collision kinetics of the system 
and the molecular interactions could result in drastically
different reaction kinetics. In the former case, unstable
subcritical nuclei have been shown to collide and coalesce
into a supercritical nucleus, but only in conditions where 
subcritical nucleation are plentiful despite a low degree
of supersaturation \cite{Baumgartner2013}. The latter case 
is more prevalent for dipole solute molecules, the anisotropic 
attraction results in nuclei with non-spherical geometries, 
this results in varying kinetics for agglomeration dependent 
on the orientation of each molecule \cite{Yau2001}. Phenomena 
such as the above examples has been collectively refereed to 
as non classical nucleation, it is known that they must be 
related to both CNT and multi-step nucleation but the connection 
is not clear based on the current literature.

As such, much of the research into nucleation theory is 
focused on developing \textit{in situ} techniques that can 
monitor and characterise the pre nucleation stage, this
would glean information on cases when both classical 
and non-classical nucleation is a possibility 
\cite{Karthika2016}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{In situ techniques for studying nucleation}
There are several different phenomena \cite{Fu2021, Karthika2016} 
that are not well described by any individual theoretical framework 
(CNT or multi-step nucleation for example). There is an increasing 
interest in probing the pre-nucleated solution.

One of the key conclusions is the need for the development of 
better \textit{in situ} techniques for characterizing and 
imaging the nucleation pathways. Understanding which specific 
pathway is unfolding under a certain set of experimental 
parameters would allow one to develop a map of the thermodynamic 
landscape available to a pre-nucleation system. The challenge 
therein lies in developing experimental methods by which one can 
reproduce similar nucleation events and in turn use 
characterisation techniques to identify the pathway taken. The 
method not only needs to be repeatable for a single set of system 
parameters, but also needs to be flexible enough to replicate 
results regardless of the system parameters (i.e. varying 
supersaturation, temperature, and solute choice). There have 
been several different methods deployed to try and observe 
crystallisation at a microscopic scale.

\subsection{Computer Simulations}
While not a direct observation of nucleation events, computer
simulations have proved an invaluable tool for studying and 
testing predictions about the pre-nucleated solution. By 
comparing the simulative nucleation rate to the expected 
rate given by current models, researchers can test newly 
developed theories and glean information about the microscopic
parameters of a system.

The main challenge facing computer simulations is the issue
of limited scope. Molecular dynamic simulations are constrained
by their choice of time and length scale, for example simulations
in a canonical ensemble face issues where the chemical potential
is depleted to the point that the driving force for crystallisation
is halted \cite{DuranOlivencia2015, Finney2023}, this can be 
addressed but requires an extensive increase in computational 
resources to maintain the chemical potential equilibria. Like wise
in most cases the choice of time scale is a crucial factor, shorter
time scales provide a more accurate evolution of the solution; 
however, in some cases the time scale magnitude for crystallisation
is often far greater than the time scale used in the simulation. This
has led to more advanced sampling techniques that allow for longer 
time scales which has provided insights into the pre-nucleation 
solution stage and formation of pre-nucleated clusters \cite{Finney2023}.

The study of the pre-nucleated solution and the evolution of 
precursors has been aided greatly by the use of simulative 
studies. Early on it was shown in colloidal crystallisation 
that the structure of the precursor was time dependent; 
colloidal simulations showed that regardless of the molecular
interactions between colloids the atomic packing pattern was 
predominately hexagonal centred, over time the packing
became dominated by body centred packing in the case 
where the interactions were longer reaching \cite{Tan2013}. 
They also confirmed that precursor formation was driven not
by local density variations, but instead was driven by the 
local bond order of the colloidal precursors \cite{Tan2013}, 
providing strong evidence that the colloidal pre-nucleation 
clusters undergo rearrangement prior to nucleation. Further
developments in $NaCl$ simulations have shown that by varying
the density fluctuations the precursor structure could either
be amorphous, the classical rock-salt structure, or a new 
structure similar to wurtzite. While this result is difficult
to confirm it provides a important conclusion, that simulations
can hint at new nucleation pathways that have not been confirmed
experimentally. This means that computer simulations can not
be merely used to confirm theories, but can be used to examine
and search for novel results.

Overall computer simulations make up a core area of research 
in the study of nucleation. The technique is however limited,
often struggling to examine nucleation over longer time scales
or larger number of particles. While sampling methods and 
enhanced simulation techniques has alleviated this somewhat 
they are still far from being able to describe the behaviour 
of larger molecules undergoing nucleation.

\subsection{Transmission Electron Microscope (TEM)}
One of the more well known techniques used for direct observations 
of crystallisation dynamics is using transmission electron 
microscopy (TEM). The basic working principle involves creating a 
focused beam of electrons that are directed onto a target sample.
Due to the wave-particle duality, the electrons can be treated 
as a unified beam of light, whose wavelength is dependent on the 
kinetic energy of the electrons. The wavelength of an individual 
electron is given by \cite{Williams2009}:
\begin{align}
	\lambda_e = \frac{h}{\sqrt{2mE(1+\frac{E}{2mc^2})}}
\end{align}

\noindent
When incident on a sample the wave is scattered by the sample 
and can then be detected by an imaging screen that measures 
the difference in intensity between the scattered and 
transmitted beams. The advantage of TEM is that one can bypass
the diffraction limit imposed by visible light, the maximum 
resolution that can be achieved by a microscope is given by
\cite{Champness2020}:
\begin{align}
	d = \frac{\lambda}{2NA}
\end{align}

\noindent
where NA is the numerical aperture of the objective lens, and
$\lambda$ is the wavelength of light used to illuminate a sample.
For a visible light LED with an average wavelength of $550~nm$, 
the upper limit for an objective lens (NA of 1.2) is around 
$230~nm$ whereas a similar lens focusing an electron beam 
with an average kinetic energy of $100~keV$ will have a maximum 
resolution of just over $4~pm$ \cite{Champness2020}. There is
an upper limit to the maximum resolution possible as electron 
lens are far inferior to what modern optics can achieve 
\cite{Champness2020} so there is a trade off between quality and 
resolution.

TEM technology has proven invaluable to the study of nucleation
at an atomic level. The samples are often only a few hundred 
atoms thick in order to optimise the imaging results. Much of 
the supporting evidence for non-classical nucleation has been 
demonstrated via TEM \cite{Ye2023}. For example when Cao 
\textit{et al} observed the formation of amorphous atomic 
cluster using $\gamma-$Fe,Au, and Re; all three of which 
underwent continuous transformation before eventually 
reaching an final ordered state \cite{Cao2020}. The amorphous 
cluster did not immediately transition to an ordered structure, 
instead fluctuating somewhat randomly until eventually settling 
into a crystalline structure \cite{Cao2020}. This was confirmed 
both in solid growth and liquid growth conditions \cite{Cao2020, 
Loh2016}, suggesting that the non-classical nucleation path 
dominates in all cases of metal crystal formation \cite{Ye2023}. 
Confirmation is necessary for different solute choices, as in 
the case of organic solids it is believed that non-classical 
nucleation only dominates in low supersaturations whereas 
classical nucleation is dominate at higher supersaturations.

TEM analysis has not been extended to the study of organic 
crystals, proteins, or simple salts at an atomic level 
however. The main reason being that organic materials can be 
damaged by the electron beam and therefore need to be frozen 
prior to being imaged \cite{Champness2020}. This was used to 
study the polymorphic time dependence of glycine crystals, by 
varying the time prior to freezing Broadhurst \textit{et al} 
showed that $\beta$ glycine dominates initially but $\alpha$-
glycine is more stable after a longer period, $\gamma$-glycine 
was only found when allowed to dry over an hour on a glass slide \cite{Broadhurst2020}. While this was invaluable for showing 
the time scales with which polymorphs form, it doesn't provide 
much information on the kinetics of crystal growth as these 
were snap shots and not direct observations of the nucleation 
process.

Overall TEM analysis is invaluable for the study of atomic 
crystal growth, the secondary scattering from electron 
beams also allows for chemical and structural information 
to be gleaned from the target. This has given strong evidence
for the non-classical nucleation theory, particularly in 
metal crystal formation. Due to the high energy of the 
electrons, organic materials have not received the same
focus outside of cryoTEM analysis. Non-invasive methods
are therefore necessary to gain an understanding of the
crystal growth and formation of organic crystals, 
particularly when both classical and multi-step nucleation 
can occur.

\subsection{Laser induced nucleation}
From as early as 1996 it has been known that laser irradiation 
using a Gaussian beam is a viable method of inducing nucleation 
within a supersaturated solution \cite{Garetz1996}. The first 
reported case was notable as it used a 1.064 $\mu m$ laser, the 
glycine solutions would appear transparent to such a laser which 
would suggest there was no photo-chemical reaction. Later studies 
into this phenomena found that the laser polarisation can influence 
the polymorph produced. With circularly polarised light producing 
$\alpha$-glycine and linearly polarised light forming $\gamma$-glycine 
\cite{Garetz2002}. Future research has found nucleation can be 
induced by 1 of 3 routes.

@@

\section{Optical Tweezers}
\subsection{Background}
Optical tweezing has been a field of applied optics ever since 
the 1980s when Ashkin \cite{Ashkin1970} first showed that focused 
light was capable of trapping micron sized particles due to light 
exerting 'radiation pressure'. The working principle was that a 
light source such as a laser could trap small objects within a 2D 
plane, as long as the light source had an approximately Gaussian 
profile (colloquially called a Gaussian beam). Soon after, Ashkin 
showed that the introduction of a microscope objective would allow 
one to focus the light source to a diffraction limited point that 
would stably trap small objects within a confined volume 
\cite{Ashkin1980}. This allowed Ashkin and others to study individual 
biological cells as the overall force was on the order of $10^{-12}\ 
N$ while being non-invasive to the internal structure. Later it would 
be used to probe microscopic properties such as the formation of 
colloidal aggregates \cite{Yi2021} to the drag forces exerted by a 
pure vacuum \cite{Ahn2018, Monteiro2018}. Due to the predictable 
behaviour of light, optical tweezers have become essential for 
measuring and exerting precise forces on the magnitude of pico-newtons 
allowing one to probe the material properties of the smallest 
materials.

\subsection{Characterisation of rotational motion}
Rotational motion about a single axis is easiest to account for.  
When the power spectra of elliptical polystyrene particles was 
fitted by Yogesh \textit{et al} \cite{Yogesha2011PreciseCO}, they 
assumed that the rotational motion was purely in the transverse
plane. As such they did not have to account for any variance in 
the trapping strength due to orientation nor did they need to
consider non-periodic rotational behaviour. In the case where 
rotational motion is stochastic the problem is more complex. 
For example, when an optical fibre trap characterisation 
technique was implemented by Saffron \textit{et al} 
\cite{BarZiv1997, Meller1998}, they were able to use dynamic 
light scattering to characterise both the axial and lateral 
trap stiffness acting on microspheres. The only drawback 
admitted to in their work was that the technique was constrained 
to isotropic scatters as their theoretical model for describing 
the auto-correlation function was predicated on the fact that 
any variations in the signal are due to the particles 
translational motion within the confines of a cylindrical trap 
\cite{BarZiv1997}. Where the upper limit of the cylindrical 
trap is given by the Rayleigh range ($z_R = n\omega_0/NA$).
However, as demonstrated by the results from Chapter 4.1, 
the axial traps of spherical aggregates is often situated 
far beyond the Rayleigh range (for a 1.2 NA laser this is 
$\pm5.985 \mu m$).

@@

However, this raises a its own host of experimental challenges,
namely how do we characterise the behaviour of an arbitrary
particle. Relying on current characterisation techniques is not
possible as they are predicated on the trapping object to behave
like an isolated sphere. Two novel methods of measuring rotational 
motion are discussed in chapter 5; firstly via a novel detection 
fibre method that allows for instantaneous measurements of the 
orientational behaviour of optically trapped ellipsoids/dimers; 
and secondly we create a simulative quadrant photo diode that 
replicates laboratory results, utilising linear regression 
techniques we measure the change in orientation in order to 
measure the optical torque applied to a non-birefringent particle.

>>>>

\chapter{Introduction}
\section{Optical Tweezers}
Precise manipulation of micro-particles and measurements of 
microscopic forces has been development in scientific research.
It has long been known that the motion of microscopic particles 
is due to multiple random collisions, measuring the average 
momentum transferred is often done by generating a potential 
well constructed using light - more commonly referred to as 
optical tweezing.

Since the mid 1900's it was known that light can transfer momentum 
to a medium, being referred to as 'radiation pressure', initially 
this was applied by transferring both linear and angular momentum 
to suspended mirrors \cite{Beth1936MechanicalDA}. This allowed 
researchers to measure the momentum of light carried by a single 
photon \cite{Beth1936MechanicalDA}. It was Ashkin who showed that 
by constructing a precise wavefront one could confine biological 
cells as small as a few microns \cite{Ashkin1970}. This was 
especially useful as the process was non-invasive, preserving the 
internal organelles while keeping the cell trapped.

The working principle is that a collimated light source - such as 
a laser - can confine a particle in the plane perpendicular to the 
beam \cite{Ashkin1970}. The radiation pressure previously observed 
is due to the momentum transferred to the target, when directed at 
an isotropic scatterer the light the momentum transfer is minimised
when the target particle is centred on the laser beam. Initially, 
confinement in the direction of the laser was achieved by matching 
the laser power so that the radiation pressure transferred would equal 
the mass of the target \cite{Ashkin1970}. Soon after, Ashkin showed 
that the by focusing the light source through a lens would confine 
the target particle in all 3 Cartesian directions regardless of the
laser power \cite{Ashkin1980}. The use of lenses to create a focal
point allowed for the development of much stronger optical traps that
could trap complex particles and biological materials.

Beyond simply observations, optical tweezing allowed for precise 
measurements of microscopic forces. Later it would be used to probe microscopic properties such as the Brownian forces exerted within a 
pure vacuum \cite{Ahn2018, Monteiro2018}. Due to the predictable 
behaviour of light, optical tweezers have become essential for 
measuring and exerting precise forces on the magnitude of pico-newtons.
However, these measurements are based on the idea that the trapped 
particle in question is an isotropic scatterer, meaning that particle 
is can be approximated as a sphere. This is an apt approximation if 
we only care about measuring the force exerted due to translational 
motion. In reality, rotational motion is a significant factor to the 
trajectory of non-spherical particles, in which case our understanding
of how light interacts with particles is incomplete without considering
the optical torque exerted by a focused beam.

\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{dimer_rotation.jpeg}
	\caption{Example of nano-dimer undergoing optical rotation 
		in a circularly polarised beam. Due to the dimer's 
		anisotropic susceptibility it is polarised along its 
		long axis, the dimer will therefore align its long axis 
		with the polarization vector. Reproduced with permission 
		from \cite{Reimann2018}.}
\end{figure}

\subsection{Characterisation of rotational motion}
Rotational motion about a single axis is easiest to account for.  
When the power spectra of elliptical polystyrene particles was 
fitted by Yogesh \textit{et al} \cite{Yogesha2011PreciseCO}, they 
assumed that the rotational motion was purely in the transverse
plane. As such they did not have to account for any variance in 
the trapping strength due to orientation nor did they need to
consider non-periodic rotational behaviour. In the case where 
rotational motion is stochastic the problem is more complex. 
For example, when an optical fibre trap characterisation 
technique was implemented by Saffron \textit{et al} 
\cite{BarZiv1997, Meller1998}, they were able to use dynamic 
light scattering to characterise both the axial and lateral 
trap stiffness acting on microspheres.

\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{BarZivDiagram.png}
	\caption{Diagram depicting the setup used by \cite{BarZiv1997}, 
	a microsphere is trapped by a focused beam. The scattered beam
	is collected by an optical fibre situated close to the trapped
	sphere. Using DLS they are able to relate the autocorrelation 
	function of the collated signal to characterise the trap stiffness 
	of the optical trap.}
\end{figure}

The only drawback admitted to in their work was that the technique 
was constrained to isotropic scatters as their theoretical model 
for describing the auto-correlation function was predicated on the 
fact that any variations in the signal are due to the particles 
translational motion within the confines of a cylindrical trap 
\cite{BarZiv1997}. Where the upper limit of the cylindrical trap 
is given by the Rayleigh range ($z_R = n\omega_0/NA$). This is 
discussed further in chapter 3 in the discussion of dimer dynamics, 
the axial traps of spherical aggregates is often situated far beyond the Rayleigh range (for a 1.2 NA laser this is $\pm5.985 \mu m$).

\section{Laser induced nucleation}
From as early as 1996 it has been known that laser irradiation 
using a Gaussian beam is a viable method of inducing nucleation 
within a supersaturated solution \cite{Garetz1996}. The first 
reported case was notable as it used a 1.064 $\mu m$ laser, the 
glycine solutions would appear transparent to such a laser which 
would suggest there was no photo-chemical reaction. Later studies 
into this phenomena found that the laser polarisation can influence 
the polymorph produced. With circularly polarised light producing 
$\alpha$-glycine and linearly polarised light forming $\gamma$-glycine 
\cite{Garetz2002}. Future research has found nucleation can be 
induced by 1 of 3 routes.

@@

@@

However, this raises a its own host of experimental challenges,
namely how do we characterise the behaviour of an arbitrary
particle. Relying on current characterisation techniques is not
possible as they are predicated on the trapping object to behave
like an isolated sphere. Two novel methods of measuring rotational 
motion are discussed in chapter 5; firstly via a novel detection 
fibre method that allows for instantaneous measurements of the 
orientational behaviour of optically trapped ellipsoids/dimers; 
and secondly we create a simulative quadrant photo diode that 
replicates laboratory results, utilising linear regression 
techniques we measure the change in orientation in order to 
measure the optical torque applied to a non-birefringent particle. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
